R for Clinical Study Reports and Submission

### 2021-09-02 tlf-ae-spec.Rmd
pass spell checking
proofread of chapter 1-8

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize which patients were included in each efficacy analysis 
in the section 12.2, adverse events of a CSR.

@@

In the code below, we count number of subjects in each AE term by SOC and treatment group.

@@

t1 %>% head(4)
```
In the code below, we count number of subjects in each AE term by SOC, AE term, and treatment group.

@@

We can also customize the text format by providing a matrix that have 
same dimention of the input dataset (i.e. `tbl_ae_spec`). 
In the code below, we illustrate how to display **bold** text for group headers
to highlight the nested structure of the table layout.

@@

>>>>

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize which patients were included in each efficacy analysis 
in Section 12.2, aAdverse Events (AEs).

@@

In the code below, we count the number of subjects in each AE term by SOC and treatment group, and we create a new variable `order` and set it as `0` for these counted numbers. 
The variable `order` will help with the data manipulation later.

@@

t1 %>% head(4)
```
In the code below, we count number of subjects in each AE term by SOC, AE term, and treatment group.
Here we also create a new variable `order` and set it as `1` for these counted numbers.

@@

We can also customize the text format by providing a matrix that have 
same dimension of the input dataset (i.e. `tbl_ae_spec`). 
In the code below, we illustrate how to display **bold** text for group headers
to highlight the nested structure of the table layout.

@@

In conclusion, the procedure to generate a specific AE summary table as shown in the above example is listed as follows:

- Step 1: Read data into R, i.e., `adae` and `adsl`.
- Step 2: Count the number of subjects by SOC and treatment group (rows with bold text) and store it in a data frame `t1`.
- Step 3: Count the number of subjects in each AE term by SOC, AE term, and treatment group (rows without bold text) and store it in a data frame `t2`.
- Step 4: Rowly bind `t1` and `t2` into `t_ae`.
- Step 5: Count the number of subjects in each arm as `t_pop`.
- Step 6: Rowly bind `t_pop` with `t_ae` as `tbl_ae_spec`.
- Step 7: Format the output by `r2rtf`.

### 2021-09-02 tlf-ae-summary.Rmd
add specific AE example
proofread of chapter 1-8

# AE Summary

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize which patients were included in each efficacy analysis 
in the section 12.2, adverse events of a CSR.

@@

>>>>

# AE Summary {#aesummary}

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize which patients were included in each efficacy analysis 
in Section 12.2, Adverse Events (AEs).

@@

In conclusion, the procedure to generate a AE summary table as shown in the above example is listed as follows:

- Step 1: Read data into R, i.e., `adae` and `adsl`.
- Step 2: Summarize participants in population by treatment group, and we name the dataset as `t_pop`.
- Step 3: Summarize participants in population by required AE criteria of interest, and we name the dataset as `t_ae`.
- Step 4: Rowly bind `t_pop` and `t_ae` and format it by `r2rtf`.

### 2021-09-02 tlf-disposition.Rmd
add analysis population example
update minor formating item
proofread of chapter 1-8

# Table for Study Patients

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize accounting of all patients who entered the study, 
in the section 10.1, disposition of patients of a CSR.

@@

>>>>

# Disposition {#disposition}

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize accounting of all patients who entered the study, in Section 10.1, Disposition of Patients.
The disposition of patients reports the numbers of patients who were randomized, and who entered and completed each phase of the study, as well as the reasons for all post-randomization discontinuations, grouped by treatment and by major reason (lost to follow-up, adverse event, poor compliance etc.).

@@

In conclusion, the procedure to generate a AE summary table as shown in the above example is listed as follows:

- Step 1: Read data into R, i.e. `adsl`.
- Step 2: Count patients in the analysis population and name the dataset as `n_rand`.
- Step 3: Count the number and percentage of patients who discontinued the study, and name the dataset as `n_disc`.
- Step 4: Count the number and percentage of of patients who discontinued the study in different reasons, and name the dataset as `n_reason`.
format it by `r2rtf`.
- Step 5: Count the number and percentage of of patients who completed the study, and name the dataset as `n_complete`.
- Step 6: Rowly bind `n_rand`, `n_disc`, `n_reason` and `n_complete` and format it by `r2rtf`.

### 2021-09-02 tlf-efficacy.Rmd
proofread of chapter 1-8

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize primary and secondary efficacy endpoints. 
in the section 11.4, efficacy results and tabulations of individual patient data of a CSR.

@@

We first define the analysis dataset using 
efficacy population flag `EFFFL` and 
all records post baseline (`AVISITN > 1`) and on or before Week 24 (`AVISITN <= 24`).

```{r}
gluc <- adlb %>%
  left_join(adsl %>% select(USUBJID, EFFFL), by = "USUBJID") %>%
  filter(EFFFL == "Y" & PARAMCD == "GLUC") %>% 
  arrange(TRTPN) %>% 
  mutate(TRTP = factor(TRTP, levels = unique(TRTP)))

@@

```{r}
ana %>%
  select(USUBJID, TRTPN, AVISIT, AVAL, BASE, CHG) %>%
  head(4)
```
## Helper Functions

@@

>>>>

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize primary and secondary efficacy endpoints,  
in Section 11.4, Efficacy Results and Tabulations of Individual Patient.

@@

We first define the analysis dataset using 
efficacy population flag `EFFFL` and 
all records post baseline (`AVISITN > 1`) and on or before Week 24 (`AVISITN <= 24`).
Here the variable `AVISITN` is the numerical analysis visit. 
For example, if the analysis visit is recorded as "Baseline", i.e., `AVISIT = Baseline`, then `AVISITN = 0`; 
if the analysis visit is recorded as "Week 24", i.e., `AVISIT = Week 24`, then `AVISITN = 24`;
if the analysis visit is blank, then `AVISITN` is also blank.
We will discuss these missing values in Section 6.4.

```{r}
gluc <- adlb %>%
  left_join(adsl %>% select(USUBJID, EFFFL), by = "USUBJID") %>%
  filter(EFFFL == "Y" & PARAMCD == "GLUC") %>% # PARAMCD is parameter code and here we focus on Glucose (mg/dL) 
  arrange(TRTPN) %>% 
  mutate(TRTP = factor(TRTP, levels = unique(TRTP)))

@@

```{r}
ana %>%
  select(USUBJID, TRTPN, AVISIT, AVAL, BASE, CHG) %>%  # AVAL: analysis value
                                                       # BASE: baseline value
                                                       # CHG: change from baseline
  head(4)
```
## Helper Functions

@@

In conclusion, the procedure to generate the above efficacy results table is summarized as follows.

- Step 1: Read the data into R, i.e., `adsl` and `adlb`.
- Step 2: Define the analysis dataset. In this example, we define two analysis datasets. The first dataset is the efficacy population (`gluc`). The second dataset is the collection of all records post baseline and on or before week 24 (`ana`).
- Step 3: Impute the missing values. In our example, we name the `ana` dataset after imputation as `ana_locf`.
- Step 4: Summarize `gluc`, i.e., calculate the mean and standard derivation, and then format it into a rtf table.
- Step 5: Summarize `ana_locf`, i.e., calculate the pairwise comparison by ANCOVA model, and then format it into a rtf table.
- Step 6: Rowly bind the output in Step 4 and Step 5.
